---
title: "bd_recreate_figures"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
Once you have a VCF you can read it in and make it a genlight object.  Follow that with making a PCA.  You will need meta data for each sample (sample name and population)

```{r}
library(tidyverse)
library(vcfR)
library(ggplot2)
library(adegenet)
library(poppr)
library(ape)
library(RColorBrewer)
library(plotly)
library(ggrepel)
library(ggforce)
library(sp)
library(sf)
library(USAboundaries)
library(rnaturalearth)
library(tmap)
library(ggmap)
library(ggpubr)
```

read in VCF and match to meta

```{r}
bd.VCF <- read.vcfR("freebayes/bd_sierra_panama_global_keep_new.vcf")

bd.meta <- read_tsv(file = "meta/bd_sierra_panama_global_meta.tsv")

# species.merge <- read_csv("meta/species_tomerge.csv")
# 
# bd.species.meta <- left_join(bd.meta, species.merge, by = "swab_id")
# write_tsv(bd.species.meta, "Rothstein_etal_Sierra_Panama_GPL_meta.txt")

colnames(bd.VCF@gt)[-1] -> vcf.names
as.data.frame(vcf.names) -> vcf.names
colnames(vcf.names) <- "swab_id"
left_join(vcf.names, bd.meta, by = "swab_id") -> vcf.meta

all(colnames(bd.VCF@gt)[-1] == vcf.meta$swab_id)

site_id_summary <- vcf.meta %>% group_by(site_id, major) %>% summarise(n=n())


```

Figure2A. Make VCF object a genlight object.  Set populations.  Make PCA.

```{r}
gl.bd <- vcfR2genlight(bd.VCF)
ploidy(gl.bd) <- 2
pop(gl.bd) <- vcf.meta$major


# all samples 
pca <- glPca(gl.bd, nf = 3)
barplot(100*pca$eig/sum(pca$eig), col = heat.colors(50), main="PCA Eigenvalues")
title(ylab="Percent of variance\nexplained", line = 2)
title(xlab="Eigenvalues", line = 1)

pca.scores <- as.data.frame(pca$scores)
pca.scores$pop <- pop(gl.bd)

pca.scores.meta<- rownames_to_column(pca.scores, var = "swab_id")

pca.scores.meta <- left_join(pca.scores.meta, vcf.meta, by = "swab_id")

pca.scores.meta <- pca.scores.meta %>% select(-c(pop, locale))

write_tsv(pca.scores.meta, "meta/pca.scores.vcf.tsv")

cols3 <- c("#000000", "#E69F00", "#56B4E9")
cols8 <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")



cols <- brewer.pal(n = nPop(gl.bd), name = "Dark2")

set.seed(9)
p.major <- ggplot(pca.scores, aes(x=PC2, y=PC1, color = pop, label = rownames(pca.scores))) + 
  geom_point(size=3) + 
  xlab("PC2 (6.6%)") +
  ylab("PC1 (21.4%)") +
  scale_color_manual(labels = c(expression(paste("Global ", italic("Bd"), "GPL")),
                     "Central Panama", 
                     "Sierra Nevada"), values = c("#E69F00","#56B4E9", "#636363")) + 
  # geom_mark_hull(aes(filter = pop == "Global")) +
  # geom_mark_hull(aes(filter = pop == "Panama")) +
  # geom_mark_hull(aes(filter = pop == "Sierra")) +
  #stat_ellipse(level = 0.95, size = 1) +
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  #scale_y_reverse() +
  theme_bw() +
  theme(legend.position = c(0.1, 0.92), text = element_text(size=18), legend.title = element_blank(), legend.background = element_rect(fill = "transparent"), legend.text.align = 0, legend.key = element_rect(fill = "transparent"))

ggplotly(p.major)

```

subset vcf by major region

```{r}
vcf.meta %>% filter(major == "Panama") -> vcf.panama.meta
vcf.meta %>% filter(major == "Sierra") -> vcf.sierra.meta

popsub(gl.bd, sublist = "Panama") -> gl.panama
popsub(gl.bd, sublist = "Sierra") -> gl.sierra

```

Figure 1C Panama PCA

```{r}
library(ggsci)
pop(gl.panama) <- vcf.panama.meta$site_id
pca.panama <- glPca(gl.panama, nf = 3)
barplot(100*pca.panama$eig/sum(pca.panama$eig), col = heat.colors(50), main="PCA Eigenvalues")
title(ylab="Percent of variance\nexplained", line = 2)
title(xlab="Eigenvalues", line = 1)

pca.panama.scores <- as.data.frame(pca.panama$scores)
pca.panama.scores$pop <- pop(gl.panama)
#pca.panama.scores$pop2 <- vcf.panama.meta$jur

cols_panama <- c('#42d4f4','#9A6324', '#4363d8', '#f58231', '#dcbeff', '#800000', '#ffe119', '#a9a9a9',  '#fabed4', '#696969')

#cols_panama <- colorRampPalette(brewer.pal(8, "Dark2"))(10)
#cols <- brewer.pal(n_distinct(pca.panama.scores$pop), "")

set.seed(9)
p.panama <- ggplot(pca.panama.scores, aes(x=PC2, y=PC1, col = pop)) + 
  geom_point(size=2) + 
  xlab("PC1 (6.0%)") + #31.0%
  ylab("PC2 (4.8%)") +   #13.8%
  #geom_mark_ellipse(aes(label = pop), tol = 0.05) +
  stat_ellipse(level = 0.95, size = 1) + 
  #scale_color_manual("black") +
  #scale_color_grey(end = 0.5) +
  scale_color_manual(values = cols_panama) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  theme_bw() +
  theme(legend.position = "none", text = element_text(size=12))
  # theme(text = element_text(size=10), legend.title = element_text(size=10),
  #         legend.text = element_text(size=10), 
  #          legend.justification=c(0,1), 
  #          legend.position=c(0.05, 0.95),
  #          legend.background = element_blank(),
  #          legend.key = element_blank())
  
```

Figure 1B Sierra Nevada PCA
```{r}
pop(gl.sierra) <- vcf.sierra.meta$site_id
pca.sierra <- glPca(gl.sierra, nf = 3)
barplot(100*pca.sierra$eig/sum(pca.sierra$eig), col = heat.colors(50), main="PCA Eigenvalues")
title(ylab="Percent of variance\nexplained", line = 2)
title(xlab="Eigenvalues", line = 1)

pca.sierra.scores <- as.data.frame(pca.sierra$scores)
pca.sierra.scores$pop <- pop(gl.sierra)
pca.sierra.scores$park <- vcf.sierra.meta$jursidiction

#cols_sierra <- brewer.pal(n_distinct(pca.sierra.scores$pop), "Paired")

cols_sierra <- c('#42d4f4','#9A6324', '#4363d8', '#f58231', '#dcbeff', '#800000', '#ffe119', '#a9a9a9',  '#fabed4', '#696969')

p.sierra <- ggplot(pca.sierra.scores, aes(x=PC2, y=PC1, col = pop)) + 
  geom_point(size=2) + 
  #geom_mark_ellipse(aes(label = pop), tol = 0.05) +
  xlab("PC2 (7.1%)") + #9.4
  ylab("PC1 (20.6%)") +   #21.3
  stat_ellipse(level = 0.95, size = 1) + 
  scale_color_manual(values = cols_sierra) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  theme_bw() +
  theme(legend.position = "none", text = element_text(size = 12))

cowplot::plot_grid(p.sierra, p.panama,  nrow = 1)
```


To make Sierra Nevada map if needed 
```{r}

# read_csv("/bd_pop_gen_combined/meta/bd_sierra_panama_only.csv") -> sierra_panama
# read_sf("../seki_summer2017/seki_meta/nps_boundary/nps_boundary.shp") %>% st_transform(crs = 4326) -> nps
# 
# sierra <- sierra_panama %>% filter(major == "sierra")
# panama <- sierra_panama %>% filter(major == "panama")
# 
# sierra_sf <- sierra %>% st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
# pana_sf  <- panama %>% st_as_sf(coords = c("longitude", "latitude"), crs = 4326)
# 
# as.factor(sierra$site_id) -> sierra$site_id
# 
# state_names <- c("california")
# park_names <- c("Kings Canyon", "Sequoia", "Yosemite")
# 
# parks <- nps %>% filter(PARKNAME %in% park_names) %>% 
#   mutate(lon=map_dbl(geometry, ~st_centroid(.x)[[1]]),
#          lat=map_dbl(geometry, ~st_centroid(.x)[[2]]))
# 
# CA<-us_states(resolution = "high", states = state_names) %>%
#   st_transform(crs = 4326)
# 
# cols_sierra <- brewer.pal(n = 10, name = "Paired")
# 
# sierra_base <- get_stamenmap(bbox = c(left = -119.5, bottom = 36.9, right = -118.3, top =38.2), zoom=11, maptype = 'terrain-labels')
# 
# sierra_map <- ggmap(sierra_base) +
#     geom_point(data = sierra, aes(x = longitude, y = latitude, col = site_id),    
#              position=position_jitter(h=0.015, w=0.015),
#              shape = 20, alpha = 0.5, size = 3) +
#   scale_color_manual(values = cols_sierra) +
#   geom_sf(data=parks, color = "gray30", lwd=0.5, fill=NA) +
#   geom_sf(data=CA, color = "gray30", lwd=0.5, fill=NA)
#   
#   
# 
# sf::write_sf(parks, "meta/parks.shp")
# 
# 
# sierra_panama %>% filter(major == "panama") %>% ggplot() + 
#   geom_point(aes(x = longitude, y = latitude, col = jursidiction), 
#              position=position_jitter(h=0.015, w=0.015),
#              shape = 20, alpha = 0.5, size = 3)
# 

```

Figure S1 Similar sample size site diversity box plots

```{r}
data_path <- "/bd_pop_gen_combined/meta/diversity_bams/theta_pestPG"
files <- dir(data_path, pattern = "*.pestPG")
file_path <- dir(data_path, pattern = "*.pestPG", full.names = T)
site <- stringr::str_split(files, pattern = "_all_out.thetas.idx.pestPG", simplify = T)
site <- site[,1]


theta_sites <- 
  map2(file_path, site, ~read_tsv(.x) %>% mutate(site_id = .y)) %>%
  reduce(rbind)

threshold <- 191
theta_sites_190 <- subset(theta_sites, theta_sites[,2] < threshold)

theta_means_sites <- theta_sites_190 %>% 
  group_by(site_id) %>% 
    summarise(
    mean_tw = mean(tW, na.rm = TRUE),
    sd_tw = sd(tW, na.rm = TRUE),
    mean_tP = mean(tP, na.rm = TRUE),
    sd_tw = sd(tP, na.rm = TRUE)
  )

## altos_de_piedra & mono_pass

res_wil_tw_alto_mono <- theta_sites_190 %>% 
  select(tW, site_id) %>% 
  filter(site_id %in% c("altos_de_piedra", "mono_pass")) %>% 
  wilcox.test(tW ~ site_id, .)

res_wil_tP_alto_mono <- theta_sites_190 %>% 
  select(tP, site_id) %>% 
  filter(site_id %in% c("altos_de_piedra", "mono_pass")) %>% 
  wilcox.test(tP ~ site_id, .)

## barrett_basin and cero negro

res_wil_tw_barrrett_cero <- theta_sites_190 %>% 
  select(tW, site_id) %>% 
  filter(site_id %in% c("barrett_basin", "cerro_negro")) %>% 
  wilcox.test(tW ~ site_id, .)

res_wil_tP_barrrett_cero <- theta_sites_190 %>% 
  select(tP, site_id) %>% 
   filter(site_id %in% c("barrett_basin", "cerro_negro")) %>% 
  wilcox.test(tP ~ site_id, .)

## unicorn_ponds & el_cope

res_wil_tw_unicorn_cope <- theta_sites_190 %>% 
  select(tW, site_id) %>% 
  filter(site_id %in% c("unicorn_ponds", "el_cope")) %>% 
  wilcox.test(tW ~ site_id, .)

res_wil_tP_unicorn_cope <- theta_sites_190 %>% 
  select(tP, site_id) %>% 
  filter(site_id %in% c("unicorn_ponds", "el_cope")) %>% 
  wilcox.test(tP ~ site_id, .)


summary(res_aov_tw_alto_mono)
summary(res_aov_tP_alto_mono)


melt_theta_sites_190 <- as.data.frame(theta_sites_190) %>% select(site_id, tW, tP) %>% 
  reshape::melt(., id.vars = "site_id", measure.vars = c("tW", "tP"))

site_id_table <- melt_theta_sites_190 %>% group_by(site_id) %>% summarise(n = n())

write_tsv(site_id_table, "/bd_pop_gen_combined/meta/diversity_bams/site_id_title.tsv")
site_id_table <- read_tsv("/bd_pop_gen_combined/meta/diversity_bams/site_id_title.tsv")

melt_theta_sites_190_size <- left_join(melt_theta_sites_190, site_id_table, by = "site_id")


write_tsv(melt_theta_sites_190, "/bd_pop_gen_combined/meta/diversity_bams/melt_thata_sites_190.tsv")

melt_theta_sites_190_size$size <- factor(melt_theta_sites_190_size$size,levels = c("N=5", "N=9", "N=12", "N=21"))


site_watter <- melt_theta_sites_190_size %>% filter(variable == "tW") %>% 
  ggplot(., aes(size, value, group = site_id, fill = region)) +
  #stat_compare_means() +
  stat_compare_means(method = "anova", label.y = 4, label = "p.signif",  bracket.size = 0.2) +
  geom_boxplot(width = 0.5) +
  ylab(expression(paste("Watterson's ", theta))) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  theme(text = element_text(size=20)) +
  scale_fill_manual(values = c("#56b4e8", "#636363")) +
  theme(legend.position = "none")

site_pi <- melt_theta_sites_190_size %>% filter(variable == "tP") %>% 
  ggplot(., aes(size, value, group = site_id, fill = region)) +
  stat_compare_means(method = "anova", label.y = 3.5, label = "p.signif",  bracket.size = 0.2) +
  geom_boxplot(width = 0.5) +
  ylab(expression(pi)) +
  theme_classic() +
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x =  element_text(size=20),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text = element_text(size = 10),
        #legend.title = element_blank(), 
        legend.title.align = 0.5) +
  scale_fill_manual(values = c("#56b4e8", "#636363"), name = "Region", labels = c("Central Panama", "Sierra Nevada"))

melt_theta_sites_190_size_sig <- melt_theta_sites_190_size %>% 
   unite(size_region, c("size", "region"))

res_aov_tW_site <- melt_theta_sites_190_size_sig %>% filter(str_detect(variable, "tW")) %>% aov(data = ., value ~ size_region)
tukey_w <- TukeyHSD(res_aov_tW_site)

res_aov_tP_site <- melt_theta_sites_190_size_sige %>% filter(str_detect(variable, "tP")) %>% aov(data = ., value ~ size_region)
tukey_p <- TukeyHSD(res_aov_tP_site)



div_plot <- cowplot::plot_grid(site_watter, site_pi + theme(legend.position = "none"), labels =  "AUTO", align = "hv")

legend <- get_legend(
  # create some space to the left of the legend
  site_pi + theme(legend.box.margin = margin(0, 0, 0, 8))
)

cowplot::plot_grid(div_plot, legend, rel_widths = c(3, .4))
  
#   
# samples.9 <- melt_theta_sites_190_size %>% 
#   filter(site_id %in% c("altos_de_piedra", "mono_pass")) %>% 
#   ggplot(., aes(variable, value,  fill= site_id)) +
#   stat_compare_means(method = "wilcox.test", label.y = 2.5) +
#   geom_boxplot(width = 0.2) +
#   scale_x_discrete(labels=c("tW" = expression(theta), "tP" = expression(pi))) +  
#   theme_classic() +
#   theme(axis.title.x = element_blank()) +
#   theme(text = element_text(size=20)) + 
#   theme(legend.position = "none") +
#   scale_fill_manual(values = c("#e69d03", "#636363")) 
# 
# samples.12 <- melt_theta_sites_190 %>% 
#   filter(site_id %in% c("barrett_basin", "cerro_negro", "conness_pond")) %>% 
#   ggplot(., aes(variable, value,  fill= site_id)) +
#   stat_compare_means(method = "wilcox.test") +
#   geom_boxplot(width = 0.2) +
#   scale_x_discrete(labels=c("tW" = expression(theta), "tP" = expression(pi))) + 
#   theme_classic() +
#   theme(axis.title.x = element_blank()) +
#   theme(text = element_text(size=20)) +
#   theme(legend.position = "none") +
#   scale_fill_manual(values = c("#636363", "#e69d03", "#636363"))
# 
# sample.21 <- melt_theta_sites_190 %>% 
#   filter(site_id %in% c("unicorn_ponds", "el_cope", "devils_crag_basin")) %>% 
#   ggplot(., aes(variable, value,  fill= site_id)) +
#   #stat_compare_means(method = "wilcox.test", label.y = 3.5) +
#   geom_boxplot(width = 0.2) +
#   scale_x_discrete(labels=c("tW" = expression(theta), "tP" = expression(pi))) + 
#   theme_classic() +
#   theme(axis.title.x = element_blank()) +
#   theme(text = element_text(size=20)) +
#   theme(legend.position = "none") +
#   scale_fill_manual(values = c("#636363", "#e69d03", "#636363")) 
# 
# samples.5 <- melt_theta_sites_190 %>% 
#   filter(site_id %in% c("leconte_divide", "el_valle_s")) %>% 
#   ggplot(., aes(variable, value,  fill= site_id)) +
#   stat_compare_means(method = "wilcox.test", label.y = 3.5) +
#   geom_boxplot(width = 0.2) +
#   scale_x_discrete(labels=c("tW" = expression(theta), "tP" = expression(pi))) + 
#   theme_classic() +
#   theme(axis.title.x = element_blank()) +
#   theme(text = element_text(size=20)) +
#   theme(legend.position = "none") +
#   scale_fill_manual(values = c("#e69d03", "#636363")) 

```

Figure 2B,C. global, panama, and sierra nevada diversity comparisions 
```{r}
global_sierra_panama <- read_tsv("global_sierra_panama_div.txt")
all_labels <- read_tsv("/bd_pop_gen_combined/div_ggplot_all_labels.txt")

global_sierra_panama <- global_sierra_panama %>% select(tW_global_noiso, tP_global_noiso, tW_panama, tP_panama, tW_sierra, tP_sierra)
melt_global_sierra_panama <- as.data.frame(global_sierra_panama) %>% 
  reshape::melt(.)

melt_global_sierra_panama_labels <- left_join(melt_global_sierra_panama, all_labels, by = "variable")
melt_global_sierra_panama_labels$region <- factor(melt_global_sierra_panama_labels$region, levels = c("Global BdGPL", "Panama", "Sierra Nevada"))

my_comparisons <- list( c("Sierra Nevada", "Panama"), c("Panama", "Global BdGPL"), c("Sierra Nevada", "Global BdGPL") )

res_aov_tW <- melt_global_sierra_panama_labels %>% filter(str_detect(variable, "tW")) %>% aov(data = ., value ~ region)
TukeyHSD(res_aov_tW)

res_aov_tP <- melt_global_sierra_panama_labels %>% filter(str_detect(variable, "tP")) %>% aov(data = ., value ~ variable)
TukeyHSD(res_aov_tP)


w_plot <- melt_global_sierra_panama_labels %>% filter(str_detect(variable,"tW")) %>% 
  ggplot(., aes(region, value, fill = region)) +
  #stat_compare_means() +
  stat_compare_means(comparisons = my_comparisons, paired = TRUE, hide.ns = TRUE, label = "p.signif") +
  geom_boxplot(width = 0.9) +
  ylab(expression(paste("Watterson's ", theta))) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(text = element_text(size=18)) +
  scale_fill_manual(values = c("#e69d03","#56b4e8", "#636363")) +
  theme(legend.position = "none")

p_plot <- melt_global_sierra_panama_labels %>% filter(str_detect(variable,"tP")) %>% 
  ggplot(., aes(region, value, fill = region)) +
  #stat_compare_means() +
  stat_compare_means(comparisons = my_comparisons, paired = TRUE, hide.ns = TRUE, label = "p.signif") +
  geom_boxplot(width = 0.9) +
  ylab(expression(pi)) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(text = element_text(size=18)) +
  scale_fill_manual(values = c("#e69d03","#56b4e8", "#636363")) +
  theme(legend.position = "none")


cow_div <- cowplot::plot_grid(w_plot, p_plot, nrow = 2, ncol = 1, labels = c("B", "C"), align = "hv")

cowplot::plot_grid(p.major, cow_div, labels = c("A"), rel_widths = c(1, 0.6), align = "hv")


```


